Building a continuous benchmarking ecosystem in bioinformatics

### 2025-02-10 review_cycle_1/highlighted_manifesto_main.tex
reviewer 2 and 3 now

@@

@@

Computational reproducibility comes from controlling the triad of data, code, and environment \cite{Hill2024-gf}. During benchmarking, a workflow runs code transforming inputs into outputs; execution is affected by the base system (OS, compiler toolchains, libraries) and its configuration. The benchmark definition should control the execution environment.

@@

\item Software dependencies cause the biggest impact on replicability and maintainability; archiving all combinations of a large dependency tree, across several OS images, exponentially increases retention and maintenance costs. Making the long-term replicability of a benchmark tractable implies adopting a sane software management system. Several solutions exist, from containerization (e.g., apptainer or singularity) \cite{Kurtzer2017-mn} and automation of reproducible and efficient software management (\texttt{EasyBuild} \cite{Hoste2012-gg} or \texttt{Spack} \cite{Gamblin2015-ll}) to distributing software installations such as CernVM-FS \cite{Blomer2013-jk},  or combinations of the two such as \texttt{computecanada} \cite{Boissonneault2019-wm} and \texttt{EESSI} \cite{Droge2023-ax}.

@@

{\color{red} Here all content on whether software is part of the backend or not, conda, pixi, web assembly etc - reviewer 3}

>>>>

{\color{red}
Aside of task completion-tailored metrics, computing performance (i.e. wallclock, on-disk read and write operations), requirements (e.g. GPUs, minimum required memory) and their scalability depending on the input size (e.g. whether runtimes increase linearly, quadratically, etc) are key, and routine, benchmarking metrics. To measure these accurately, benchmarking systems ideally would provide means to: i, characterize the hardware running the benchmark, so methods runtimes are comparable if run on homogeneous systems;  ii, track performances (CPU or GPU usage, memory) when running each benchmark component; iii, introduce little overhead during the profiling (so profiling does not considerably slow down runtimes); iv, collect variability in performance by running each method several times; v, go beyond reporting average resources usage and provide them over execution time (e.g. so embarrasingly parallel algorithms can be told apart from largely single-cored runs with few parallelism bursts).
}

@@

{\color{red}
Mature communities are key in sustaining and reflecting the existing challenge in a given field. For instance, CASP, a continuous benchmarking by nature, has a established community defining the critical assesments scope, input data, target definitions, and composite and normalized metrics to equally weight multiple performance scores \cite{simpkin2023tertiary}. At the continuous benchmark level, CASP keeps track of past performances and introduces metrics to evaluate possible artifacts, i.e. whether round progress could be arising from different target difficulty \cite{kryshtafovych2019critical}. To facilitate community contribution, we believe benchmarking systems can lower the barrier to contributing, hence making the contribution efforts less likely to be biased. 
}

@@

Computational reproducibility comes from controlling the triad of data, code, and environment \cite{Hill2024-gf}. During benchmarking, a workflow runs code transforming inputs into outputs; execution is affected by the base system ({\color{red}\sout{OS} operating system}, compiler toolchains, libraries) and its configuration. The benchmark definition should control the execution environment.

@@

\item Software dependencies cause the biggest impact on replicability and maintainability; archiving all combinations of a large dependency tree, across several OS images, exponentially increases retention and maintenance costs. Making the long-term replicability of a benchmark tractable implies adopting a sane software management system. Several solutions exist, from containerization (e.g., apptainer or singularity) \cite{Kurtzer2017-mn} and automation of reproducible and efficient software management (\texttt{EasyBuild} \cite{Hoste2012-gg} or \texttt{Spack} \cite{Gamblin2015-ll}) to distributing software installations such as CernVM-FS \cite{Blomer2013-jk},  or combinations of the two such as the Digital Research Alliance of Canada (former \texttt{computecanada}) \cite{Boissonneault2019-wm} and \texttt{EESSI} \cite{Droge2023-ax}.

@@

{\color{red} 
The responsibility of handling the software environment poses an interesting challenge to benchmarking systems. On one hand, some workflow managers provide means to install and activate software and re-run workflows when software specification changes, even if the workflow topology remains the same. For these, software is the workflow manager's responsability. On the other hand, some benchmarks might be addressed to comparing the same workflow topology run with different software backends (i.e., with conda vs compilled) or with different versions of some software dependencies components (i.e. several samtools/htslib releases). Benchmarking systems should provide enough flexibility to take control and make the latter possible.

Finally, scientific software management handling is known for its nuances, including abundant dependencies with known incompatibilities and complex installation recipes \cite{droge2023-ax}. Scientific software management is handled via traditional (linux) package managers (Ri.e. PMs, DEBs), central software stacks and enviornment modules (i.e. Lmod), conda, containers, or read-only network filesystems providing ready-to-use installations (CernVM-FS). These strategies are under active development: EESSI \cite{droge2023-ax} and the Digital Research Alliance of Canada \cite{Boissonneault2019-wm} are effectively providing multi-purpose, reusable and efficient software installations via CernVM-FS; conda is speeding up their deployment procedures (via py-rattler/pixi) and allowing compilation to Web assembly (to execute code within Web browser)s; or apptainer (former Singularity) provides rootless containarization suitable to run in scientific clusters. 
}
